---
title: "Introduction to bestSelectR"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to bestSelectR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(bestSelectR)
```

## Overview

**bestSelectR** performs best subset selection for logistic regression models. It evaluates all possible combinations of predictor variables and identifies the subsets that best predict a binary outcome.

This vignette demonstrates the basic functionality and shows you how to:

- Perform basic subset selection
- Interpret the results
- Use different selection metrics
- Make predictions with fitted models
- Handle missing data

## Basic Example

Let's start with a simple example using the built-in `mtcars` dataset to predict transmission type (automatic vs manual):

```{r basic-example}
# Prepare data
data(mtcars)
X <- as.matrix(mtcars[, c("mpg", "hp", "wt", "qsec")])
y <- mtcars$am # 0 = automatic, 1 = manual

# Find best subsets
result <- bestSubset(X, y, max_variables = 3, top_n = 5, metric = "auc")

# View results
print(result)
```

The output shows:
- Best model information (which variables, performance metrics)
- Selection settings used
- Data summary

## Understanding the Results

Use `summary()` to see detailed information:

```{r summary}
summary(result)
```

This shows:
- Top 5 models ranked by AUC
- Coefficients of the best model
- Performance metrics for each model

## Extracting Information

You can extract specific components:

```{r extract-info}
# Get coefficients of best model
coeffs <- coef(result)
print(coeffs)

# Access best model details
best <- result$best_model
cat("Best model uses", best$n_variables, "variables\n")
cat("Variables:", paste(colnames(X)[best$variables], collapse = ", "), "\n")
cat("AUC:", round(best$auc, 3), "\n")

# View all top models
top_models <- result$models
head(top_models)
```

## Making Predictions

Use the fitted model to predict on new data:

```{r predictions}
# Create new data (same structure as training data)
new_data <- matrix(c(
  20.0, 150, 3.0, 18.0, # Car 1
  15.0, 300, 4.5, 16.0, # Car 2
  25.0, 100, 2.5, 20.0 # Car 3
), nrow = 3, ncol = 4, byrow = TRUE)
colnames(new_data) <- colnames(X)

# Probability predictions
probs <- predict(result, new_data, type = "prob")
cat("Predicted probabilities:\n")
print(probs)

# Class predictions
classes <- predict(result, new_data, type = "class")
cat("\nPredicted classes:\n")
print(classes)
```

## Selection Metrics

Different metrics emphasize different aspects of model quality:

### AUC (Default)

Area Under the ROC Curve - measures discrimination ability:

```{r auc-metric}
result_auc <- bestSubset(X, y, max_variables = 3, metric = "auc")
cat("Best AUC:", round(result_auc$best_model$auc, 3), "\n")
```

### Accuracy

Proportion of correct predictions:

```{r accuracy-metric}
result_acc <- bestSubset(X, y, max_variables = 3, metric = "accuracy")
cat("Best accuracy:", round(result_acc$best_model$accuracy, 3), "\n")
```

### BIC

Bayesian Information Criterion - penalizes model complexity:

```{r bic-metric}
result_bic <- bestSubset(X, y, max_variables = 3, metric = "bic")
cat("Best BIC:", round(result_bic$best_model$bic, 2), "\n")
```

**Metric guidance:**
- **AUC**: Good for imbalanced data, measures discrimination
- **Accuracy**: Simple, intuitive for balanced classes
- **BIC/AIC**: Statistical criteria, penalize complexity
- **Deviance**: Direct measure of model fit (lower is better)

## Cross-Validation

Use cross-validation for more reliable performance estimates:

```{r cross-validation}
result_cv <- bestSubset(X, y,
  max_variables = 3,
  cross_validation = TRUE,
  cv_folds = 5,
  cv_seed = 123,
  metric = "auc"
)

cat("Cross-validated AUC:", round(result_cv$best_model$auc, 3), "\n")
```

Cross-validation helps avoid overfitting by testing the model on data it hasn't seen.

## Handling Missing Data

The package provides three approaches to missing values:

```{r missing-data, error=TRUE}
# Create data with missing values
X_missing <- X
X_missing[c(1, 5, 10), 1] <- NA

# Option 1: Error on missing values (default)
try(bestSubset(X_missing, y, na.action = na.fail))

# Option 2: Remove rows with missing values
result_omit <- bestSubset(X_missing, y, na.action = na.omit, max_variables = 2)
cat("Used", result_omit$call_info$n_observations, "complete cases\n")

# Option 3: Remove rows but preserve row alignment
result_exclude <- bestSubset(X_missing, y, na.action = na.exclude, max_variables = 2)
```

## Performance Tips

For large datasets or many predictors:

1. **Limit maximum variables**: Use `max_variables` to reduce computation time
2. **Use parallel processing**: Set `n_threads` for faster evaluation
3. **Consider the computational cost**: The package evaluates all possible subsets up to `max_variables`

```{r performance-tips, eval=FALSE}
# Limit to 5 variables for speed
result_fast <- bestSubset(X, y, max_variables = 5)

# Use multiple threads (auto-detected)
result_parallel <- bestSubset(X, y, max_variables = 8, n_threads = NULL)

# Specify thread count
result_threads <- bestSubset(X, y, max_variables = 8, n_threads = 4)
```

**Computational complexity:**

- For `max_variables = k` and `p` predictors: evaluates C(p,1) + C(p,2) + ... + C(p,k) models
- Example: `p = 20, max_variables = 8` evaluates ~264,000 models
- Recommendation: Keep `max_variables ≤ 20` for reasonable speed

## Complete Example

Here's a complete workflow:

```{r complete-example}
# 1. Prepare data
set.seed(123)
X <- matrix(rnorm(100), nrow = 20, ncol = 5)
colnames(X) <- paste0("X", 1:5)
y <- rbinom(20, 1, 0.5)

# 2. Find best subsets with cross-validation
result <- bestSubset(X, y,
  max_variables = 3,
  top_n = 3,
  metric = "auc",
  cross_validation = TRUE,
  cv_folds = 5,
  cv_seed = 42
)

# 3. Examine results
summary(result)

# 4. Extract best model coefficients
coef(result)

# 5. Make predictions on new data
new_X <- matrix(rnorm(25), nrow = 5, ncol = 5)
colnames(new_X) <- paste0("X", 1:5)
predictions <- predict(result, new_X, type = "prob")
cat("\nPredictions for new data:\n")
print(predictions)
```

## Additional Resources

- Package documentation: `?bestSubset`
- README with test plan: See package GitHub repository
- Method implementations: `?predict.bestSubset`, `?coef.bestSubset`, `?summary.bestSubset`

## Summary

Key takeaways:

- `bestSubset()` finds optimal variable combinations for logistic regression
- Use `max_variables` to control computational cost
- Different metrics (`auc`, `accuracy`, `bic`) emphasize different model qualities
- Cross-validation provides more reliable performance estimates
- The package handles missing data with `na.action` parameter
- Use `predict()` for new data, `coef()` for coefficients, `summary()` for detailed results
